---
title: "Day 8 Practice"
output: html_document
---

1. Write an if...else statement to test whether a string matches "right", and if so, print "you win!", but if the string matches "wrong", print "you lose!", and if it's neither, print "the result is inconclusive".

```{r}

string <- "none"

if (string == "right") {
  print("you win!")
} else if (string == "wrong") {
  print("you lose!")
} else {
  print ("the result is inconclusive")
}

```

2. Make a copy of the iris data. Use an ifelse statement to create a new column, where:
  - if the value of Sepal.Length is less than 5, the new column should have a value of "low"
  - if the value of Sepal.Length is greater than 6, it should have a value of "high"
  - and if it's between 5 and 6 (inclusive), it should have a value of "mid"
  - make a plot in ggplot2 to confirm that you did this right
  
```{r}

myiris <- iris

myiris$High_Low <- 
  ifelse(myiris$Sepal.Length < 5, "low", 
  ifelse(myiris$Sepal.Length > 6, "high", "mid"))


library(ggplot2)

ggplot(myiris, (aes(x = Sepal.Length , y = High_Low))) +
  geom_point()



```

3. Consider the iris data again.
  - make a scatterplot of Sepal.Length by Sepal.Width
  - facet the plot by Species
  - using a loop, loop through species and create three separate plots
    - in each plot, display the species as the plot title 
      (check `?labs` for how to set the plot title in a convenient way)
    - use pdf() and dev.off() so that the three plots get written out to a file
    
```{r}
head(myiris)

ggplot(myiris, aes(x = Sepal.Length, y = Sepal.Width)) +
  geom_point() +
  facet_grid (~ Species)

pdf("class stuff")
for (i in unique(myiris$Species)) {
  print(ggplot(myiris[myiris$Species == i, ], aes(x = Sepal.Length, y = Sepal.Width)) +
  geom_point())
  labs(title = i)
} 
dev.off()

# get() # Pull the object that has that string name


```

4. Similar to what we did with the correlation power simulations, write code (including a loop) that creates a power simulation for comparing the means of two normally-distributed sample.
  - compare the sample with both `t.test()` and `wilcox.test()`
  - put the p-values from both tests in each simulation result
  - compare the power of the t-test vs. the power of the Wilcoxon test
  
```{r}
x <- rnorm(1000, mean = 5, sd = 2)
summary(x)
sd(x)

y <- rnorm(1000, mean = 10, sd = 2)
summary(y)
sd(y)

t.test(x, y)

summary$Mean(rnorm(100, mean = 5, sd = 2))

str(mean(rnorm(100, mean = 5, sd = 2)))

for ttest in

t.test(samples.x[1,], samples.y[1,])

for (i in 1:100) {
  samples.x <- data.frame(i, rnorm(i, mean = 5, sd = 2))
}

samples.x

samples.x

for (j in 1:100) {
  samples.y <- rnorm(j, mean = 10, sd = 2)
}




```

