---
title: "Day 7 Homework"
output: html_document
---

Find three relationships in your data.  For each relationship, do the following:

1. Verbally articulate the relationship you want to examine.

Relationship #1 
What is the relationship between Homicide Rates and Time

Relationship #2 
What is the relationship between the Percentage of the Population between 15 and 24 years and time

Relationship #2 
What is the relationship between Homicide Rates and the Percentage of the Population between 15 and 24 years


2. Verbally identify what variables you want to map to each graphical dimension.

Relationship #1 - HomicideRate & Year
Relationship #2 - Pop15_24 & Time
Relationship #3 - HomicideRate & Pop15_24

3. Reshape your data as needed.

```{r}
library(dplyr)
library(ggplot2)

#I am loading a dataset that I am currently formatting, that includes only Homicides. The code for this data is in the "WHO Mortality.Rmd" file
load("data/Homicide.Cause.RData")
load("data/Population.RData")

#The unit of analysis is the number of deaths by cause, sex, year and country. I am collapsing it to country and year. I am just keepign some variables
Homicide.Sum <- data.frame(
  summarize(group_by(Homicide.Cause, Country, Year), 
            Homicides = sum(Deaths1)))

head(Population)
Population.Sum <- data.frame(
  summarize(group_by(Population, Country, Year), 
            Population = sum(Pop1),
            Pop15_24 = sum(Pop9, Pop10)))

#Merging Population Data
Homicide.Pop <- merge(Homicide.Sum, Population.Sum, all = TRUE)

#Generate New Variables

Homicide.Pop$HomicideRate <- (Homicide.Pop$Homicides/Homicide.Pop$Population) * 100000

Homicide.Pop$Per15_24 <- (Homicide.Pop$Pop15_24/Homicide.Pop$Population) * 100

head(Homicide.Pop)

save(Homicide.Pop, file = "Data/Homicide.Pop.RData")


```


4. Use ggplot2 to create your graphs.

Relationship #1 - HomicideRate & Year

```{r}
mean1 <- summarize(group_by(Homicide.Pop, Year), mean(HomicideRate, na.rm=TRUE))

ggplot(mean1, aes(Year, `mean(HomicideRate, na.rm = TRUE)`)) +
  geom_line()

```

Relationship #2 - Per15_24 & Time

```{r}
mean2 <- summarize(group_by(Homicide.Pop, Year), mean(Per15_24, na.rm=TRUE))
mean2

ggplot(mean2, aes(Year, `mean(Per15_24, na.rm = TRUE)`)) +
  geom_line()

```


Relationship #3 - HomicideRate & Pop15_24

```{r}
ggplot(Homicide.Pop, aes(Per15_24, HomicideRate)) +
  geom_point()

mean3 <- merge(mean1, mean2, na.rm=TRUE)

ggplot(mean3, aes(Year)) +
  geom_line(aes(y = `mean(Per15_24, na.rm = TRUE)`)) +
  geom_line(aes(y = `mean(HomicideRate, na.rm = TRUE)`))


```


5. Fiddle with colors, axis labels, themes, etc. until the plot looks "publication ready" to you.  Try to change at least one thing per plot.

```{r}
library(scales)

#Scatter-Plot
ggplot(Homicide.Pop, aes(Per15_24, HomicideRate)) +
  geom_point(alpha = .5) +
  geom_smooth(method = lm, color = "red") +
  theme_bw () +
  theme(text = element_text(family = "serif", size = 12),
        panel.border = element_blank()) +
  scale_y_log10(breaks = seq(0, 500, 50)) +
  xlab("Percent 15 to 24") +
  ylab("Homicide Rate")

#Line Graph
#Selecting Just the United States
Homicide.Pop[Homicide.Pop$Country == 2450,]

#Question: I have been tying to plot these two line graphs with two y scales, one in the right, but I could not find a way to do that.

ggplot(Homicide.Pop[Homicide.Pop$Country == 2450,], aes(Year)) +
  geom_line(aes(y = HomicideRate)) +
  geom_line(aes(y = Per15_24))

ggplot(Homicide.Pop[Homicide.Pop$Country == 2450,], aes(x = Year)) +
  geom_line(aes(y = HomicideRate, color = "red"), size = 1.2) +
  geom_line(aes(y = Per15_24, color = "blue"), size = 1.2) +
  ylab("Homicide Rate                               Percent 15 to 24") +
  theme_bw () +
  theme(legend.position = "none", 
        text = element_text(family = "serif", size = 12),
        panel.border = element_blank())

#Question: Use annotate based on a variable. Mark the ICD version of each piece of the series


```


6. Save all plots to a format that you could use in a paper.  HINT: PNG works well for pasting into Word documents.

```{r}
png("day07_homework_dots.png")
ggplot(Homicide.Pop, aes(Per15_24, HomicideRate)) +
  geom_point(alpha = .5) +
  geom_smooth(method = lm, color = "red") +
  theme_bw () +
  theme(text = element_text(family = "serif", size = 12),
        panel.border = element_blank()) +
  scale_y_log10(breaks = seq(0, 500, 50)) +
  xlab("Percent 15 to 24") +
  ylab("Homicide Rate")
dev.off()

png("day07_homework_line.png")


ggplot(Homicide.Pop[Homicide.Pop$Country == 2450,], aes(x = Year)) +
  geom_line(aes(y = HomicideRate, color = "red"), size = 1.2) +
  geom_line(aes(y = Per15_24, color = "blue"), size = 1.2) +
  ylab("Homicide Rate                               Percent 15 to 24") +
  theme_bw () +
  theme(legend.position = "none", 
        text = element_text(family = "serif", size = 12),
        panel.border = element_blank())

dev.off()
```

