---
title: "Day 4 Homework"
output: html_document
---

1. Find some aspect of your data that needs to be combined or merged in some way.  This could be a variety of things:
  - multiple "sheets" in an Excel
  - multiple files that should be combined into one data set
  - properties that need to be merged in to one data set. For example:
    - you have a table of subject properties (gender, handedness, etc.) that need to be merged into the item-level data
    - you have other information on countries (population, etc.) that need to be merged in
  - if you DON'T have this naturally-occurring issue, figure out a way to manufacture it. For example:
    - split up your data into chunks
    - make up some new table of information that you want to merge in
  - Explain (in words) what you are trying to do
  - Use binding/merging commands to accomplish this

Answer 1.
  -My data is divided in 5 different zip files. 
  -I am attempting to combine all these files in a single data frame.
    -Including an id variable that identifies the data source (ICD)

```{r}
library('dplyr')

Mortality <- bind_rows("10_2" = read.csv(unzip("Data/Morticd10_part2.zip")),
                       "10_1" = read.csv(unzip("Data/Morticd10_part1.zip")),
                       "9" = read.csv(unzip("Data/morticd9.zip")),
                       "8" = read.csv(unzip("Data/morticd08.zip")),
                       "7" = read.csv(unzip("Data/morticd07.zip")),
                       .id = "ICD")
save (Mortality, file = "Data/Mortality.RData")

Population <- read.csv(unzip("Data/Pop.zip"))
save(Population, file = "Data/Population.RData")
```

I had two issues:
  -Issue #1: "ICD" was generated as a character variable, instead of a factor
    -I found no solution within the bind_rows function. Is there one?
  -Issue #2: My working folder was littered with large files for each piece from the file extraction.
  -The best solutions I found are bellow:

Issue #1
  -Is there a way to prevent the variable from being transformed?

```{r}
Mortality$ICD <- factor(as.factor(Mortality$ICD), c("7","8","9","10_1","10_2"))
summary (Mortality$ICD)

```

Issue #2
  -Question: Is there a way to prevent these files from being generated? I am researching about temp files. Is there a   better solution?

```{r}
file.remove("Morticd10_part2","Morticd10_part1","Morticd9","Morticd8","MortIcd7","pop")

```

