---
title: "Day 3 Practice"
output: html_document
---

1. Make a "copy" of the `iris` data set in your workspace. Use `summary()` on it, identify a factor column.

```{r}
myiris <- iris
summary(myiris)
summary(myiris$Species)

```

2. Create different versions of the factor column, in the new data set:
   a. a numeric vector with `as.numeric()`
   b. a character vector
   
```{r}

myiris$numspecies <- as.numeric(myiris$Species)
myiris$charspecies <- as.character(myiris$Species)
summary(myiris)


```

3. Use `xtabs()` to see how the values in these new columns line up with the original factor column.

```{r}
xtabs(~myiris$Species + numspecies, myiris)
xtabs(~myiris$Species + charspecies, myiris)

#All three
xtabs(~myiris$Species + numspecies + charspecies, myiris)

```

4. Check the order of levels of Species with `levels()`. Create a new column from this factor, changing the order of the levels, using:
  a. `relevel()`
  b. `reorder()`
  c. `factor()`

```{r}

summary(myiris)
summary(myiris$Species)
levels(myiris$Species)

#a
myiris$SpeciesRel <- relevel(myiris$Species,"virginica")
summary(myiris$SpeciesRel)

#b
myiris$SpeciesReo <- reorder(levels(myiris$Species),c(2,1,3))
summary(myiris$SpeciesReo)

#c
myiris$SpeciesFac <- factor(myiris$Species, levels =  c("virginica","setosa","versicolor"))
summary(myiris$SpeciesFac)

```

5. Working with missing values
   - read in the data from "CognitionPaperModifiedData.csv"
   - locate NAs
   - display the rows of the data frame where the WithinMinPairs variable has missing values
   - create a new column that's a copy of the WithinMinPairs column
   - replace all missing values in this column with zeros
   #Ignore maybe. Why it does not help me? Missing data - use `xtabs()` to check this change in coding

```{r}
scottdata <- read.csv("other/CognitionPaperModifiedData.csv")

# locate NAs
summary(scottdata)

#display the rows of the data frame where the WithinMinPairs variable has missing values
scottdata[scottdata$WithinMinPairs == NA, ]
scottdata[is.na(scottdata$WithinMinPairs), ]

#create a new column that's a copy of the WithinMinPairs column
scottdata$WithinMinPairsCopy <- scottdata$WithinMinPairs

#replace all missing values in this column with zeros

  #Does not work
scottdata$WithinMinPairsCopy[scottdata$WithinMinPairsCopy == NA] <- 0
summary (scottdata$WithinMinPairsCopy)

  #Works
scottdata$WithinMinPairsCopy[is.na(scottdata$WithinMinPairsCopy)] <- 0
summary (scottdata$WithinMinPairsCopy)

#use `xtabs()` to check this change in coding
t <- xtabs(~WithinMinPairs + WithinMinPairsCopy, scottdata, na.action = na.pass, exclude = NULL)

str(t)

t[c(1,2,3),c(1,2,231)]


```

