---
title: "Final Project_MSantos"
author: "Mateus Renno Santos"
date: "January 22, 2017"
output: html_document
---

## Setup
Clean the workspace, and installs the 'usePackage' function.

```{r Setup}
#Clear the workspace
rm(list = ls())

# Function - usePackage()
  #Library a package, and installs the package if necessary. Use this function instead of "library()"
usePackage <- function(x) {
    if (!is.element(x, installed.packages()[,1]))
        install.packages(x, dep = TRUE)
    require(x, character.only = TRUE)
}

usePackage('dplyr')
usePackage('ggplot2')
usePackage('scales')
usePackage('gridExtra')

```


## Extract Data
Generates the Mortality and Population dataframes from the files downloaded from the WHO Mortality Database Website, to a "Data" folder.
URL: http://www.who.int/healthinfo/statistics/mortality_rawdata/en/

```{r Extract}
usePackage('dplyr')

Mortality <- bind_rows("10_2" = read.csv(unz("Data/Morticd10_part2.zip","Morticd10_part2")),
                       "10_1" = read.csv(unz("Data/Morticd10_part1.zip","Morticd10_part1")),
                       "9" = read.csv(unz("Data/morticd9.zip","Morticd9")),
                       "8" = read.csv(unz("Data/morticd08.zip","Morticd8")),
                       "7" = read.csv(unz("Data/morticd07.zip","MortIcd7")),
                       .id = "ICD")

Mortality$ICD <- factor(as.factor(Mortality$ICD), c("10_2","10_1","9","8","7"))

Population <- read.csv(unz("Data/Pop.zip","pop"))

save (Mortality, file = "Data/Mortality.RData")
save (Population, file = "Data/Population.RData")


```


## Homicide
Removes all non-homicide causes of deaths from the Mortality dataframe. Generates the Homicide dataframe

```{r Homicide}

Homicide <- Mortality[c(which(Mortality$List=="07A" & Mortality$Cause=="A149"),
                        which(Mortality$List=="07B" & Mortality$Cause=="B050"),
                        which(Mortality$List=="08A" & Mortality$Cause=="A148"),
                        which(Mortality$List=="08B" & Mortality$Cause=="B050"),
                        which(Mortality$List=="09A" & Mortality$Cause=="B55"),
                        which(Mortality$List=="09B" & Mortality$Cause=="B55"),
                        which(Mortality$List=="09C" & Mortality$Cause=="C103"),
                        which(Mortality$List=="09N" & (Mortality$Cause=="B55" | Mortality$Cause=="CH17")),
                        which(Mortality$List=="101" & (Mortality$Cause=="1102" | Mortality$Cause=="1103")),
                        which(Mortality$List=="103" & (Mortality$Cause=="X85" | Mortality$Cause=="X86" | 
                                                    Mortality$Cause=="X87" | Mortality$Cause=="X88" | 
                                                    Mortality$Cause=="X89" | Mortality$Cause=="X90" | 
                                                    Mortality$Cause=="X91" | Mortality$Cause=="X92" | 
                                                    Mortality$Cause=="X93" | Mortality$Cause=="X94" | 
                                                    Mortality$Cause=="X95" | Mortality$Cause=="X96" | 
                                                    Mortality$Cause=="X97" | Mortality$Cause=="X98" | 
                                                    Mortality$Cause=="X99" | Mortality$Cause=="Y00" | 
                                                    Mortality$Cause=="Y01" | Mortality$Cause=="Y02" | 
                                                    Mortality$Cause=="Y03" | Mortality$Cause=="Y04" | 
                                                    Mortality$Cause=="Y05" | Mortality$Cause=="Y06" | 
                                                    Mortality$Cause=="Y07" | Mortality$Cause=="Y08" | 
                                                    Mortality$Cause=="Y09")),
                        which(Mortality$List=="104" & (Mortality$Cause=="X850" | Mortality$Cause=="X851" | 
                                                    Mortality$Cause=="X852"	| Mortality$Cause=="X853"	| 
                                                    Mortality$Cause=="X854" | Mortality$Cause=="X855" | 
                                                    Mortality$Cause=="X856" | Mortality$Cause=="X857" | 
                                                    Mortality$Cause=="X858" | Mortality$Cause=="X859" | 
                                                    Mortality$Cause=="X860" | Mortality$Cause=="X861" | 
                                                    Mortality$Cause=="X862" | Mortality$Cause=="X864" | 
                                                    Mortality$Cause=="X865" | Mortality$Cause=="X866" | 
                                                    Mortality$Cause=="X868" | Mortality$Cause=="X869" | 
                                                    Mortality$Cause=="X870" | Mortality$Cause=="X871" | 
                                                    Mortality$Cause=="X872" | Mortality$Cause=="X874" | 
                                                    Mortality$Cause=="X875" | Mortality$Cause=="X876" | 
                                                    Mortality$Cause=="X877" | Mortality$Cause=="X878" | 
                                                    Mortality$Cause=="X879" | Mortality$Cause=="X880" | 
                                                    Mortality$Cause=="X881" | Mortality$Cause=="X882" | 
                                                    Mortality$Cause=="X883" | Mortality$Cause=="X884" | 
                                                    Mortality$Cause=="X885" | Mortality$Cause=="X886" | 
                                                    Mortality$Cause=="X887" | Mortality$Cause=="X888" | 
                                                    Mortality$Cause=="X889" | Mortality$Cause=="X890" | 
                                                    Mortality$Cause=="X891" | Mortality$Cause=="X892" | 
                                                    Mortality$Cause=="X893" | Mortality$Cause=="X894" | 
                                                    Mortality$Cause=="X895" | Mortality$Cause=="X896" | 
                                                    Mortality$Cause=="X897" | Mortality$Cause=="X898" | 
                                                    Mortality$Cause=="X899" | Mortality$Cause=="X900" | 
                                                    Mortality$Cause=="X901" | Mortality$Cause=="X902" | 
                                                    Mortality$Cause=="X903" | Mortality$Cause=="X904" | 
                                                    Mortality$Cause=="X905" | Mortality$Cause=="X906" | 
                                                    Mortality$Cause=="X907" | Mortality$Cause=="X908" | 
                                                    Mortality$Cause=="X909" | Mortality$Cause=="X910" | 
                                                    Mortality$Cause=="X911" | Mortality$Cause=="X912" | 
                                                    Mortality$Cause=="X913" | Mortality$Cause=="X914" | 
                                                    Mortality$Cause=="X915" | Mortality$Cause=="X916" | 
                                                    Mortality$Cause=="X917" | Mortality$Cause=="X918" | 
                                                    Mortality$Cause=="X919" | Mortality$Cause=="X920" | 
                                                    Mortality$Cause=="X921" | Mortality$Cause=="X922" | 
                                                    Mortality$Cause=="X923" | Mortality$Cause=="X924" | 
                                                    Mortality$Cause=="X925" | Mortality$Cause=="X926" | 
                                                    Mortality$Cause=="X927" | Mortality$Cause=="X928" | 
                                                    Mortality$Cause=="X929" | Mortality$Cause=="X930" | 
                                                    Mortality$Cause=="X931" | Mortality$Cause=="X932" | 
                                                    Mortality$Cause=="X933" | Mortality$Cause=="X934" | 
                                                    Mortality$Cause=="X935" | Mortality$Cause=="X936" | 
                                                    Mortality$Cause=="X937" | Mortality$Cause=="X938" | 
                                                    Mortality$Cause=="X939" | Mortality$Cause=="X940" | 
                                                    Mortality$Cause=="X941" | Mortality$Cause=="X942" | 
                                                    Mortality$Cause=="X943" | Mortality$Cause=="X944" | 
                                                    Mortality$Cause=="X945" | Mortality$Cause=="X946" | 
                                                    Mortality$Cause=="X947" | Mortality$Cause=="X948" | 
                                                    Mortality$Cause=="X949" | Mortality$Cause=="X950" | 
                                                    Mortality$Cause=="X951" | Mortality$Cause=="X952" | 
                                                    Mortality$Cause=="X953" | Mortality$Cause=="X954" | 
                                                    Mortality$Cause=="X955" | Mortality$Cause=="X956" | 
                                                    Mortality$Cause=="X957" | Mortality$Cause=="X958" | 
                                                    Mortality$Cause=="X959" | Mortality$Cause=="X960" | 
                                                    Mortality$Cause=="X961" | Mortality$Cause=="X962" | 
                                                    Mortality$Cause=="X963" | Mortality$Cause=="X964" | 
                                                    Mortality$Cause=="X965" | Mortality$Cause=="X966" | 
                                                    Mortality$Cause=="X967" | Mortality$Cause=="X968" | 
                                                    Mortality$Cause=="X969" | Mortality$Cause=="X970" | 
                                                    Mortality$Cause=="X971" | Mortality$Cause=="X972" | 
                                                    Mortality$Cause=="X973" | Mortality$Cause=="X974" | 
                                                    Mortality$Cause=="X975" | Mortality$Cause=="X976" | 
                                                    Mortality$Cause=="X977" | Mortality$Cause=="X978" | 
                                                    Mortality$Cause=="X979" | Mortality$Cause=="X980" | 
                                                    Mortality$Cause=="X981" | Mortality$Cause=="X982" | 
                                                    Mortality$Cause=="X983" | Mortality$Cause=="X984" | 
                                                    Mortality$Cause=="X985" | Mortality$Cause=="X987" | 
                                                    Mortality$Cause=="X988" | Mortality$Cause=="X989" | 
                                                    Mortality$Cause=="X990" | Mortality$Cause=="X991" | 
                                                    Mortality$Cause=="X992" | Mortality$Cause=="X993" | 
                                                    Mortality$Cause=="X994" | Mortality$Cause=="X995" | 
                                                    Mortality$Cause=="X996" | Mortality$Cause=="X997" | 
                                                    Mortality$Cause=="X998" | Mortality$Cause=="X999" | 
                                                    Mortality$Cause=="Y000" | Mortality$Cause=="Y001" | 
                                                    Mortality$Cause=="Y002" | Mortality$Cause=="Y003" | 
                                                    Mortality$Cause=="Y004" | Mortality$Cause=="Y005" | 
                                                    Mortality$Cause=="Y006" | Mortality$Cause=="Y007" | 
                                                    Mortality$Cause=="Y008" | Mortality$Cause=="Y009" | 
                                                    Mortality$Cause=="Y010" | Mortality$Cause=="Y011" | 
                                                    Mortality$Cause=="Y012" | Mortality$Cause=="Y013" | 
                                                    Mortality$Cause=="Y014" | Mortality$Cause=="Y015" | 
                                                    Mortality$Cause=="Y016" | Mortality$Cause=="Y017" | 
                                                    Mortality$Cause=="Y018" | Mortality$Cause=="Y019" | 
                                                    Mortality$Cause=="Y020" | Mortality$Cause=="Y021" | 
                                                    Mortality$Cause=="Y022" | Mortality$Cause=="Y023" | 
                                                    Mortality$Cause=="Y024" | Mortality$Cause=="Y025" | 
                                                    Mortality$Cause=="Y026" | Mortality$Cause=="Y027" | 
                                                    Mortality$Cause=="Y028" | Mortality$Cause=="Y029" | 
                                                    Mortality$Cause=="Y030" | Mortality$Cause=="Y031" | 
                                                    Mortality$Cause=="Y032" | Mortality$Cause=="Y033" | 
                                                    Mortality$Cause=="Y034" | Mortality$Cause=="Y035" | 
                                                    Mortality$Cause=="Y036" | Mortality$Cause=="Y037" | 
                                                    Mortality$Cause=="Y038" | Mortality$Cause=="Y039" | 
                                                    Mortality$Cause=="Y040" | Mortality$Cause=="Y041" | 
                                                    Mortality$Cause=="Y042" | Mortality$Cause=="Y043" | 
                                                    Mortality$Cause=="Y044" | Mortality$Cause=="Y045" | 
                                                    Mortality$Cause=="Y046" | Mortality$Cause=="Y047" | 
                                                    Mortality$Cause=="Y048" | Mortality$Cause=="Y049" | 
                                                    Mortality$Cause=="Y050" | Mortality$Cause=="Y051" | 
                                                    Mortality$Cause=="Y052" | Mortality$Cause=="Y053" | 
                                                    Mortality$Cause=="Y054" | Mortality$Cause=="Y055" | 
                                                    Mortality$Cause=="Y056" | Mortality$Cause=="Y057" | 
                                                    Mortality$Cause=="Y058" | Mortality$Cause=="Y059" | 
                                                    Mortality$Cause=="Y060" | Mortality$Cause=="Y061" | 
                                                    Mortality$Cause=="Y062" | Mortality$Cause=="Y068" | 
                                                    Mortality$Cause=="Y069" | Mortality$Cause=="Y070" | 
                                                    Mortality$Cause=="Y071" | Mortality$Cause=="Y072" | 
                                                    Mortality$Cause=="Y073" | Mortality$Cause=="Y078" | 
                                                    Mortality$Cause=="Y079" | Mortality$Cause=="Y080" | 
                                                    Mortality$Cause=="Y081" | Mortality$Cause=="Y082" | 
                                                    Mortality$Cause=="Y083" | Mortality$Cause=="Y084" | 
                                                    Mortality$Cause=="Y085" | Mortality$Cause=="Y086" | 
                                                    Mortality$Cause=="Y087" | Mortality$Cause=="Y088" | 
                                                    Mortality$Cause=="Y089" | Mortality$Cause=="Y090" | 
                                                    Mortality$Cause=="Y091" | Mortality$Cause=="Y092" | 
                                                    Mortality$Cause=="Y093" | Mortality$Cause=="Y094" | 
                                                    Mortality$Cause=="Y095" | Mortality$Cause=="Y096" | 
                                                    Mortality$Cause=="Y097" | Mortality$Cause=="Y098" | 
                                                    Mortality$Cause=="Y099" | Mortality$Cause=="Y871")),
                        which(Mortality$List=="10M" & (Mortality$Cause=="Y871" | Mortality$Cause=="X85" | 
                                                    Mortality$Cause=="X86" | Mortality$Cause=="X87" | 
                                                    Mortality$Cause=="X88" | Mortality$Cause=="X89" | 
                                                    Mortality$Cause=="X90" | Mortality$Cause=="X91" | 
                                                    Mortality$Cause=="X92" | Mortality$Cause=="X93" | 
                                                    Mortality$Cause=="X94" | Mortality$Cause=="X95" | 
                                                    Mortality$Cause=="X96" | Mortality$Cause=="X97" | 
                                                    Mortality$Cause=="Y02" | Mortality$Cause=="Y03" | 
                                                    Mortality$Cause=="Y04" | Mortality$Cause=="Y05" | 
                                                    Mortality$Cause=="Y06" | Mortality$Cause=="Y07" | 
                                                    Mortality$Cause=="Y08" | Mortality$Cause=="Y09")),
                        which(Mortality$List=="UE1" & Mortality$Cause=="UE64")), ]

save (Homicide, file = "Data/Homicide.RData")

```


## Reshape
Change the unit of analysis of the Mortality and Population dataframes from "Cause", by "Sex", by "Year", by "Country", to "Cause", by "Year", by "Country". The "Sex" collumn gets reshaped to a wide format.

```{r Reshape}
#Converts the Sex collumns to a named factor.

Homicide$Sex <- factor(Homicide$Sex, 
                        levels = c(1, 2, 9), 
                        labels = c("M", "F", "U"))

Population$Sex <- factor(Population$Sex, 
                        levels = c(1, 2, 9), 
                        labels = c("M", "F", "U"))


#Verify for missing data in any ID variables
for (i in c(1:7, 9:10)) {
  print(colnames(Homicide[i]))
  print(table(is.na(Homicide[,i])))
}

for (i in c(1:4, 6)) {
  print(colnames(Population[i]))
  print(table(is.na(Population[,i])))
}

#Converts the missing values in the ID variables "Admin1" & "SubDiv" to "Country". Missing data in those variables refer to country level data.

Homicide$Admin1[is.na(Homicide$Admin1)] <- "Country"
Homicide$SubDiv[is.na(Homicide$SubDiv)] <- "Country"
Homicide$SubDiv[Homicide$SubDiv == ""] <- "Country"

Population$Admin1[is.na(Population$Admin1)] <- "Country"
Population$SubDiv <- as.character(Population$SubDiv)
Population$SubDiv[is.na(Population$SubDiv)] <- "Country"
Population$SubDiv[Population$SubDiv == ""] <- "Country"

#Reshape
Homicide <- reshape(Homicide, 
                    timevar = "Sex", 
                    idvar = colnames(Homicide[,c(1:7, 9:10)]), 
                    direction = "wide")


```


## Collapse
Aggregates all "Causes"" of Homicides in a single number by "Country", and "Year". 

```{r Collapse}

Homicide <- data.frame(
  summarise_all(group_by(Homicide[,-7], 
                         Country, Year, Admin1, SubDiv, ICD, List, Frmat, IM_Frmat),
                funs(sum), na.rm = TRUE))

```


## Totals
Generates the total homicides collumns for each "Country", "Year", and age group.

```{r Totals}

for (i in 1:26) {
  Homicide[paste("Deaths", i, sep = "")] <-
    apply(Homicide[, c(i+8,i+38,i+68)], MARGIN = 1, FUN = sum, na.rm = TRUE)
}

for (i in 1:4) {
  Homicide[paste("IM_Deaths", i, sep = "")] <-
    apply(Homicide[, c(i+34,i+64,i+94)], MARGIN = 1, FUN = sum, na.rm = TRUE)
}

```


## Population
Reshape, Generate Totals and Merge the Population dataframe with the Homicide dataframe

```{r Population}

# Reshape
Population <- reshape(Population, 
                      timevar = "Sex", 
                      idvar = colnames(Population[,c(1:4, 6)]),
                      direction = "wide")

# Totals
for (i in 1:26) {
  Population[paste("Pop", i, sep = "")] <-
    apply(Population[, c(i+5,i+32,i+59)], MARGIN = 1, FUN = sum, na.rm = TRUE)
}

Population["Lb"] <-
    apply(Population[, c(32, 59, 86)], MARGIN = 1, FUN = sum, na.rm = TRUE)

names(Population)[names(Population) == 'Frmat'] <- 'FrmatPop'

# Merge
HomPop <- merge(Homicide, Population, all = TRUE)


```


## Labels
Generates a "Name" variable for country names. Directly sourced from the WHO Country Codes table.

```{r Labels}

Codes <- read.csv(unz("Data/country_codes.zip","country_codes"))

HomPop$Name <- factor(HomPop$Country, levels = Codes$country, labels = Codes$name)

#Moves the "Name" to the first collumn of the df
usePackage('dplyr')

HomPop <- select(HomPop, Name, everything())

```


#Rates
Generates homicide rates and other variables from transformations

```{r Rates}

#Homicide Rate
HomPop$HomRate <- (HomPop$Deaths1/HomPop$Pop1) * 100000

#Age Distribution
HomPop$Per15_24 <- ((HomPop$Pop9 + HomPop$Pop10)/HomPop$Pop1)*100


save(HomPop, Homicide, Mortality, Population, file = "Data/Objects Jan22_2017.RData")

#load("Data/Objects Jan22_2017.RData")

```


# Aggregated Trends
Plots for all data points with valid data in the dataframe

```{r Plots All}
usePackage('ggplot2')
usePackage('scales')

# 1. Investigate the prevalence of missing Homicide and Population Data by Country
  #Year; Deaths1; Pop1

  #Homicide
ggplot(HomPop[HomPop$Admin == "Country",], aes(x = Year)) +
    geom_point(aes(y = Deaths1), size = 1.5, color = "dark red", alpha = .5) +
    ggtitle("Total Homicides by Year") +
    ylab("Total Number of Homicides") +
    xlab("Year") +
    scale_x_continuous(breaks=seq(1950, 2015, 5), limits = c(1950, 2015)) +
    scale_y_continuous(labels = comma) +
    theme(
      text = element_text(family = "serif", size = 12),
      panel.grid.major.x = element_line(colour = "white"),
      panel.grid.major.y = element_line(colour = "gray"),
      panel.background = element_rect(fill = "white"))

  #Population
ggplot(HomPop[HomPop$Admin == "Country",], aes(x = Year)) +
    geom_point(aes(y = Pop1), size = 1.5, color = "dark red", alpha = .5) +
    ggtitle("Total Population by Year") +
    ylab("Total Population") +
    xlab("Year") +
    scale_x_continuous(breaks=seq(1950, 2015, 5), limits = c(1950, 2015)) +
    scale_y_continuous(labels = comma) +
    theme(
      text = element_text(family = "serif", size = 12),
      panel.grid.major.x = element_line(colour = "white"),
      panel.grid.major.y = element_line(colour = "gray"),
      panel.background = element_rect(fill = "white"))

# 2. Investigate shared trends in Homicide Rate and in the Proportion of the Population between 15 to 24 years
  #Year; HomRate; Per15_24

  #Homicide Rate
ggplot(HomPop[HomPop$Admin == "Country",], aes(x = Year)) +
    geom_point(aes(y = HomRate), size = 1.5, color = "dark red", alpha = .5) +
    ggtitle("Homicide Rate by Year") +
    ylab("Homicide Rate per 100,000 Population") +
    xlab("Year") +
    scale_x_continuous(breaks=seq(1950, 2015, 5), limits = c(1950, 2015)) +
    scale_y_continuous(labels = comma) +
    theme(
      text = element_text(family = "serif", size = 12),
      panel.grid.major.x = element_line(colour = "white"),
      panel.grid.major.y = element_line(colour = "gray"),
      panel.background = element_rect(fill = "white"))

  #Population 15 to 24
ggplot(HomPop[HomPop$Admin == "Country",], aes(x = Year)) +
    geom_point(aes(y = Per15_24), size = 1.5, color = "dark red", alpha = .5) +
    ggtitle("Percent of the Population Aged 15 to 24 per Year") +
    ylab("Percent 15 to 24") +
    xlab("Year") +
    scale_x_continuous(breaks=seq(1950, 2015, 5), limits = c(1950, 2015)) +
    scale_y_continuous(labels = comma) +
    theme(
      text = element_text(family = "serif", size = 12),
      panel.grid.major.x = element_line(colour = "white"),
      panel.grid.major.y = element_line(colour = "gray"),
      panel.background = element_rect(fill = "white"))

# 3. Investigate the Association between Homicide Rates and the Proportion of the Population between 15 to 24 years
  #HomRate; Per15_24

ggplot(HomPop[HomPop$Admin == "Country", ], aes(Per15_24/100)) +
  geom_smooth(aes(y = HomRate), method = "lm") +
  geom_point(aes(y = HomRate), alpha = .5) +
  ggtitle("Population 15 to 24 and Homicide Rate by Year") +
  ylab("Homicide Rate per 100,000 population") +
  xlab("Percent of the Population Aged 15 to 24") +
  scale_x_continuous(labels = percent) +
  scale_y_log10(labels = comma) +
  theme(
    text = element_text(family = "serif", size = 12),
    panel.grid.major.x = element_line(colour = "white"),
    panel.grid.major.y = element_line(colour = "gray"),
    panel.background = element_rect(fill = "white"))

```


## Country-level Trends
Investigate trends in the data with individualized graphs by country

```{r Plots Countries, message=FALSE, warning=FALSE}
usePackage('ggplot2')
usePackage('scales')
usePackage('gridExtra')

dir.create("Plots", showWarnings = FALSE)

# 1. Investigate the prevalence of missing Homicide and Population Data by Country
  #Deaths1 & Pop1

pdf("Plots/Year_Hom_Pop.pdf")

for (i in unique(HomPop$Country)) {
  
  Top <- 
    ggplot(HomPop[HomPop$Country == i & HomPop$Admin == "Country", ], aes(x = Year)) +
    geom_line(aes(y = Deaths1), size = .5, color = "red") +
    geom_point(aes(y = Deaths1), size = 1.5, color = "dark red") +
    ggtitle(paste("Total Homicides - ",
                  unique(HomPop$Name[HomPop$Country == i]), " (",
                  unique(HomPop$Country[HomPop$Country == i]), ")")) +
    ylab("Total Number of Homicides") +
    xlab("Year") +
    scale_x_continuous(breaks=seq(1950, 2015, 5), limits = c(1950, 2015)) +
    scale_y_continuous(labels = comma) +
    theme(
      text = element_text(family = "serif", size = 12),
      panel.grid.major.x = element_line(colour = "white"),
      panel.grid.major.y = element_line(colour = "gray"),
      panel.background = element_rect(fill = "white"))

  Bottom <- 
    ggplot(HomPop[HomPop$Country == i & HomPop$Admin == "Country", ], aes(x = Year)) +
    geom_line(aes(y = Pop1), size = .5, color = "blue") +
    geom_point(aes(y = Pop1), size = 1.5, color = "dark blue") +
    ggtitle(paste("Total Population - ",
                  unique(HomPop$Name[HomPop$Country == i]), " (",
                  unique(HomPop$Country[HomPop$Country == i]), ")")) +
    xlab("Year") +
    ylab("Total Population") +
    scale_x_continuous(breaks=seq(1950, 2015, 5), limits = c(1950, 2015)) +
    scale_y_continuous(labels = comma) +
    theme(
      text = element_text(family = "serif", size = 12),
      panel.grid.major.x = element_line(colour = "white"),
      panel.grid.major.y = element_line(colour = "gray"),
      panel.background = element_rect(fill = "white"))
  
  grid.arrange(Top, Bottom)
  
  print(paste(unique(HomPop$Name[HomPop$Country == i]), "(",
              unique(HomPop$Country[HomPop$Country == i]), ") Done."), sep = NULL)
  
}

dev.off()

# 2. Investigate shared trends in Homicide Rate and in the Proportion of the Population between 15 to 24 years
  #Year; HomRate; Per15_24
     
pdf("Plots/Year_HomRate_Per1524.pdf")

for (i in unique(HomPop$Country)) {
  
  Top <- 
    ggplot(HomPop[HomPop$Country == i & HomPop$Admin == "Country", ], aes(x = Year)) +
    geom_line(aes(y = HomRate), size = .5, color = "red") +
    geom_point(aes(y = HomRate), size = 1.5, color = "dark red") +
    ggtitle(paste("Homicide Rate - ",
                  unique(HomPop$Name[HomPop$Country == i]), " (",
                  unique(HomPop$Country[HomPop$Country == i]), ")")) +
    ylab("Homicide Rate per 100,000 population") +
    xlab("Year") +
    scale_x_continuous(breaks=seq(1950, 2015, 5), limits = c(1950, 2015)) +
    scale_y_continuous(labels = comma) +
    theme(
      text = element_text(family = "serif", size = 12),
      panel.grid.major.x = element_line(colour = "white"),
      panel.grid.major.y = element_line(colour = "gray"),
      panel.background = element_rect(fill = "white"))

  
  Bottom <- 
    ggplot(HomPop[HomPop$Country == i & HomPop$Admin == "Country", ], aes(x = Year)) +
    geom_line(aes(y = Per15_24), size = .5, color = "blue") +
    geom_point(aes(y = Per15_24), size = 1.5, color = "dark blue") +
    ggtitle(paste("Percent Aged 15 to 24 - ",
                  unique(HomPop$Name[HomPop$Country == i]), " (",
                  unique(HomPop$Country[HomPop$Country == i]), ")")) +
    xlab("Year") +
    ylab("Percent of the Population Aged 15 to 24") +
    scale_x_continuous(breaks=seq(1950, 2015, 5), limits = c(1950, 2015)) +
    scale_y_continuous(labels = percent) +
    theme(
      text = element_text(family = "serif", size = 12),
      panel.grid.major.x = element_line(colour = "white"),
      panel.grid.major.y = element_line(colour = "gray"),
      panel.background = element_rect(fill = "white"))
  
  grid.arrange(Top, Bottom)
  
  print(paste(unique(HomPop$Name[HomPop$Country == i]), "(",
              unique(HomPop$Country[HomPop$Country == i]), ") Done."), sep = NULL)

}

dev.off()

# 3. Investigate the Association between Homicide Rates and the Proportion of the Population between 15 to 24 years
  #HomRate; Per15_24

pdf("Plots/Scatter_HomRate_Per1524.pdf")

for (i in unique(HomPop$Country)) {
  dots <- ggplot(HomPop[HomPop$Country == i & HomPop$Admin == "Country", ], 
                 aes(x = (Per15_24/100))) +
    geom_smooth(aes(y = HomRate), method = "lm") +
    geom_point(aes(y = HomRate), alpha = .5) +
    ggtitle(paste("Population 15 to 24 and Homicide Rate - ",
                  unique(HomPop$Name[HomPop$Country == i]), " (",
                  unique(HomPop$Country[HomPop$Country == i]), ")")) +
    ylab("Homicide Rate per 100,000 population") +
    xlab("Percent of the Population Aged 15 to 24") +
    scale_x_continuous(labels = percent) +
    scale_y_continuous(labels = comma) +
    theme(
      text = element_text(family = "serif", size = 12),
      panel.grid.major.x = element_line(colour = "white"),
      panel.grid.major.y = element_line(colour = "gray"),
      panel.background = element_rect(fill = "white"))
  
  grid.arrange(dots)

  print(paste(unique(HomPop$Name[HomPop$Country == i]), "(",
              unique(HomPop$Country[HomPop$Country == i]), ") Done."), sep = NULL)

}

dev.off()

```


# Models
Summary models of the relationship between the percentage of the population between 14 and 24 years of age, and the homicide rate of countries. OLS estimates. More adequade fixed effect models require adjsments to the data, but will be executed in the future.

```{r Models}

# Simple Linear Regression with an Unstransformed Dependent Variable
ols <-lm(HomRate ~ Per15_24, data=HomPop)

summary(ols)

ols$coefficients[2]

# However, the distribution of Homicide Rates is highly skewed to the right.
hist(HomPop$HomRate)

# Since there are no real zeros in the scale of homicide rates, I will log-transform the dependent variable

HomPop$Log_HomRate <- log(HomPop$HomRate)
HomPop$Log_HomRate[which(is.infinite(HomPop$Log_HomRate))] <- NA

ols_log <-lm(Log_HomRate ~ Per15_24, data=HomPop, na.action = na.omit)

summary(ols_log)

exp(ols_log$coefficients[2])

```

On average, each addition percentage point in the proportion of the population between 15 to 24 years of age, corresponds to `r exp(ols_log$coefficients[2])` more homicides. Results are preliminary. They only consider countries and years with observable valid numbers. The model also desconsider the dependence in residuals within countries.

I will later execute a fixed effects model (plm), but only after complementing and cleaning the data.

In particular, many rates are missing due to lack of data on population size, which can be easily acessible from other sources.
